The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-04 analysis/paper/paper.qmd
Update paper

For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage. Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated. As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words. <!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant imoprvement was observed in the result. --> Subject-specific stop words, we manually constructed a list based on our domain knowledge. Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well. Finally, titles and author names were removed as document stop words on document-level.

@@

@@

emu_theses_words %>% 
  mutate(stem = wordStem(words = word, 
                         language = "porter")) |>  # Implement the Porter stemming algorithm for English provided by the SnowballC package
  count(stem, sort = TRUE)
```

@@

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the beta statistic. The probability of a document being part of a topic is given by the gamma statistic.

```{r}
# The {topicmodels} package requires a document-term matrix as input
emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
emu_theses_dtm

# Fit LDA model
emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
emu_theses_lda

# Tidy model with beta statistic
emu_theses_topics <- tidy(emu_theses_lda, matrix = "beta")
emu_theses_topics

# The top n words that are the most common in each topic
emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)

emu_theses_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# 
beta_spread <- emu_theses_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

beta_spread

beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  slice_max(abs(log_ratio), n = 20) %>% 
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(log_ratio, term)) +
  geom_col() +
  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)

emu_theses_documents <- tidy(emu_theses_lda, matrix = "gamma")
emu_theses_documents

emu_theses_documents |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip()

emu_theses_documents |> 
  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_theses_documents |> 
  inner_join(emu_theses |> select(title, graduation_year), by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", stem = TRUE)

spacyr::spacy_parse(emu_theses_bigrams$w_1)

bigram_counts <- emu_theses_bigrams |> 
  count(w_1, w_2, sort = TRUE)

bigram_stem_counts <- emu_theses_bigrams |> 
  count(w_1_stem, w_2_stem, sort = TRUE)

bigram_counts
bigram_stem_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text")

trigram_counts <- emu_theses_trigrams |> 
  count(w_1, w_2, w_3, sort = TRUE)

>>>>

For analyses in which word-level tokenisation was chosen, stop words, i.e., common words or words that do not add much meaning, were removed in the data preparation stage. Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated. As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words. <!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant improvement was observed in the result. --> We manually constructed a list of subject-specific stop words based on our domain knowledge. Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well. Finally, titles and author names were removed as document stop words.

@@

# TODO can we use this instead of the short_words function?
emu_theses_words <- emu_theses_words |> 
  filter(nchar(word) > 4)

@@

emu_theses_words %>% 
  mutate(stem = wordStem(words = word, 
                         language = "porter")) |>  # Implement the Porter stemming algorithm for English provided by the SnowballC package
  count(stem, sort = TRUE)

@@

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.

```{r topic-model-lda}

# Prepare document-term matrix as input for LDA model
emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")

# Create LDA model
emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))

```{r topic-model-beta}

# Extract topics as mixtures of terms from LDA model
emu_topics <- tidy(emu_lda, matrix = "beta")

# The top n most common words in each topic
emu_theses_top_terms <- get_top_words_per_topic(emu_topics, 20)

# TODO create visualisation function
emu_theses_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# TODO can this be generalised to more than two topics?
beta_spread <- emu_theses_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  slice_max(abs(log_ratio), n = 20) %>% 
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(log_ratio, term)) +
  geom_col() +
  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
```

```{r topic-model-gamma}

# Extract documents as mixtures of topics from LDA model
emu_documents <- tidy(emu_lda, matrix = "gamma")

emu_documents |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip()

emu_documents |> 
  inner_join(emu_theses |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = reorder(str_sub(document, 1, 50), gamma), y = gamma, fill = topic)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  facet_wrap(~ graduation_year, ncol = 1, scales = "free", drop = TRUE)

emu_documents |> 
  inner_join(emu_theses |> 
               select(title, graduation_year), 
             by = c("document" = "title")) |> 
  ggplot(aes(x = factor(graduation_year), y = topic, fill = factor(topic))) +
  geom_col()
```

@@

## TO DO: generalise bigrams and trigrams into ngrams function

## Bigrams
emu_theses_bigrams <- emu_theses |> 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |> 
  separate(bigram, into = c("first", "second"), sep = " ", remove = FALSE) |> 
  anti_join(stop_words, by = c("first" = "word")) |> 
  anti_join(stop_words, by = c("second" = "word")) |> 
  filter(str_detect(first, "[a-z]") & 
         str_detect(second, "[a-z]"))

bigram_counts <- emu_theses_bigrams |> 
  count(first, second, sort = TRUE)

bigram_counts

## Trigrams
emu_theses_trigrams <- emu_theses |> 
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
  separate(trigram, into = c("first", "second", "third"), sep = " ", remove = FALSE) |> 
  filter(!first %in% stop_words$word,
         !second %in% stop_words$word,
         !third %in% stop_words$word) |>  
  filter(str_detect(first, "[a-z]"), 
         str_detect(second, "[a-z]"),
         str_detect(third, "[a-z]"))

trigram_counts <- emu_theses_trigrams |> 
  count(first, second, sort = TRUE)
